---
title: "Group8 Draft"
date: "2023-03-14"
output: 
  pdf_document:
          number_sections: yes
fig_caption: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(visdat)
library(gridExtra)
library(tinytex)
```

## Summary



```{r}
dataset8 <- read_csv("dataset8.csv")
summary(dataset8)
```
## Missing data
```{r dataset8}
vis_miss(dataset8)
```
## Plots

IMDB movie Characteristics

```{r echo=TRUE}
p1 <- ggplot(data=dataset8, aes(x=genre)) + 
       geom_bar(fill="lightblue") + 
       xlab("Genre") +
       ylab("Movies") +
       theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
p2 <- ggplot(data=dataset8, aes(x=year)) + 
       geom_bar(fill="lightblue") + 
       xlab("Year") +
       ylab("Movies") +
       theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
p3<- ggplot(data=dataset8, aes(x=length)) + 
  geom_bar(fill="lightblue") + 
  xlab("Length") +
  ylab("Movies") +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0))
p4 <- ggplot(data=dataset8, aes(x=rating)) + 
  geom_histogram(binwidth=5, fill="lightblue") +
  geom_vline(xintercept=mean(dataset8$rating), colour='red', linetype='longdash') +
  geom_text(label='Mean', x=55, y=60, hjust='center', size=3) +
  xlab("Rating")
p5 <- ggplot(data=dataset8, aes(x=budget)) + 
  geom_histogram(binwidth=10, fill="lightblue") +
  xlab("Budget")
p6 <- ggplot(data=dataset8, aes(x=votes)) + 
  geom_histogram(binwidth=10, fill="lightblue") +
  xlab("Votes")
grid.arrange(p2, p3, p1, p4, p5, p6, nrow=4,
             top="Movie Characteristics")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


```{r boxplots}
imdb_group_08 <- dataset8 %>%
  mutate(Rating_above_7 = ifelse(rating > 7,"Yes","No"))
# Convert chr to factor
imdb_group_08$Rating_above_7 <- as.factor(imdb_group_08$Rating_above_7)
imdb_group_08$genre <- as.factor(imdb_group_08$genre)

p7 <- ggplot(data = imdb_group_08, mapping = aes(x = Rating_above_7, y = year)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "if rating is greater than 7", y = "year",
       title = "Binary Boxplot of Year by Rating")  +
  theme(plot.title = element_text(size=7))
p8 <- ggplot(data = imdb_group_08, mapping = aes(x = Rating_above_7, y = length)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "if rating is greater than 7", y = "length",
       title = "Binary Boxplot of Length by Rating")  +
  theme(plot.title = element_text(size=7))
p9 <- ggplot(data = imdb_group_08, mapping = aes(x = Rating_above_7, y = budget)) +
  geom_boxplot(fill = "steelblue") +
  labs(x = "if rating is greater than 7", y = "budget",
       title = "Binary Boxplot of Budget by Rating")  +
  theme(plot.title = element_text(size=7))
grid.arrange(p7, p8, p9, ncol=3)
```
# Creating an additional column that indicate yes or no if rating is greater than 7

##Modeling 
predictor included, by rating variable, using generalized liner model
fit model
if binomial, 0<y<1, the variabale rating not in.
```{r model}
model<-glm(rating~year+length+budget+votes,data=dataset8)
#summary(model)
```
