---
title: "The Impact of firm's features on IMDB Ratings"
author: "Yujie Tang, Jialu FU, Weiqing GUO, Bashiru Mukaila, Wanding Wang"
date: "2023-03-14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(MASS)
library(tidyverse)
library(kableExtra)
library(broom)
library(sjPlot)
library(stats)
library(jtools)
library(GGally)
library(ggplot2)
```

# Introduction

The entertainment industry happens to be a competitive market and the film industry is a subset of it. The stakeholders in the film industry are interested in features that make their films successful. One key aspect of a film's success is its rating on platforms such as IMDB, which can influence audience perception and drive revenue.

The aim of this project is to investigate the relationship between the properties of films and their IMDB ratings. Specifically, we want to use a logistic regression model to know which properties of a firm influence whether a film is rated by IMDB as greater than 7 or not, using variables such as year of release, length, budget, number of votes, and genre.

```{r expl1, echo=FALSE}
# Importing the data set
group_08 <- read.csv("C:/Users/mbashiru/OneDrive - AIICO INSURANCE/Desktop/MSc. Data Analytics/Second Semester/STATS5085 - Data Analysis Skills (M)/Assessment/4. Group Project 2/dataset8.csv")

# Creating an additional column that indicate yes or no if rating is greater than 7
imdb_group_08 <- group_08 %>%
  mutate(Rating_above_7 = ifelse(rating > 7,"Yes","No"))
```

```{r expl2, echo=FALSE, include=FALSE}
# Displaying the structure of the data set
str(imdb_group_08)
```

# Exploratory Data Analysis

You can also embed plots, for example:

```{r expl3, echo=FALSE}
# Displaying the summary statistics of the numerical variables
summary(imdb_group_08[c("year", "length", "budget", "votes", "rating")])
```


```{r scatp, echo=FALSE}
# Create a scatterplot matrix of the numerical variables
ggpairs(imdb_group_08[c("year", "length", "budget", "votes", "rating")])
```

```{r boxp, echo=FALSE}
# Creating a box plot of the IMDB ratings by genre
ggplot(imdb_group_08, aes(x = genre, y = rating)) + 
  geom_boxplot()
```

